---
title: "Video Game Sales and Performance Analysis"
#author: "Sairam Venkatachalam"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: lumen
    css: |
      body {
        background-color: #0074e4; /* Blue background */
        color: #ffffff; /* White text */
      }
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session.
library(ezids)
library(ggplot2)
library(dplyr)
```


# 1. Introduction


## 1.1 Project Topic

The central focus of this project revolves around the comprehensive examination of video game sales and performance, delving into their dynamics, trends, and underlying factors to uncover valuable insights. Being avid video game enthusiasts, our team is looking to deep dive into the history of video game sales in order to find answers to our research questions

## 1.2 Data Source

The dataset used for this report was sourced through web scraping, specifically from Metacritic. It is important to note that certain observations are missing, as Metacritic covers only a subset of the gaming platforms. As a result, some games may not have complete information for all the variables. However, the dataset contains approximately 6,900 complete cases.

The dataset includes the following fields:

* Name
* Platform
* Year_of_Release
* Genre
* Publisher
* NA_Sales
* EU_Sales
* JP_Sales
* Other_Sales
* Global_Sales
* Critic_score (an aggregate score by Metacritic staff)
* Critic_count (the number of critics used for the Critic_score)
* User_score (score by Metacritic's subscribers)
* User_count (the number of users who provided the user_score)
* Developer (the party responsible for creating the game)
* Rating (ESRB ratings)


## 1.3 Project objectives and research questions

We will commence our analysis by conducting Exploratory Data Analysis on the dataset to unveil intriguing trends and patterns within the data.

After the completion of EDA, we wish to find answers to the following research questions:

* What proportion of shooter genre video games features a user score exceeding 8.0?
* What are the average user scores for games developed by Nintendo, and how do they compare with those from other game developers? Are these differences statistically significant?
* What is the trend in Global Sales over the years, and how has the recency of platforms influenced their popularity?
* Is there a statistically meaningful difference in mean sales between games with a rating of 7.5 or higher and those with a rating below 7.5? Can we identify a critical score associated with increased sales?



```{r}
games<-read.csv('Video games sales.csv')
games<-na.omit(games)
games$Rating <- ifelse(games$Rating== "", NA,games$Rating )
games<-subset(games,games$Rating!="NA")
games<-subset(games,games$Year_of_Release!="N/A")

```


# 4. Research Questions


```{r}

games_final= outlierKD2(games,Global_Sales,rm=TRUE,hist=FALSE)
games_final<-na.omit(games_final)

```

## Research Question 1: 

**What proportion of all 'Shooter' Games have a User Score of >=8.0**

The idea of this research question was to get a sense of the proportion or percentage of games in the genre of 'Shooter' which had a high user rating which we defined at 8.0 or more. This was done in the following steps:

* First, we calculate the point estimate for the proportion from our sample
* Second, we calculate the standard error for out sample distribution. We use the t distribution as we don't know the standard deviation of the overall population
* Third, we calculate the critical t score that captures the 95% confidence interval for the proportion
* Finally, we calculate the upper and lower bounds using the point estimate and standard errors

```{r}
#----------------------------------------------------------------------------------------------------------------
# SMART Q - Confidence interval for Shooter game ratings

subset_data <- subset(games_final,games_final$Genre == 'Shooter' & games_final$User_Score >= 8.0)
shooter_data <- subset(games_final,games_final$Genre == 'Shooter')


sample_proportion <- nrow(subset_data) / nrow(shooter_data)

confidence_level <- 0.95

standard_error <- sqrt((sample_proportion * (1 - sample_proportion)) / nrow(shooter_data))

df <- nrow(shooter_data) - 1  # Degrees of freedom
t_critical <- qt(1 - (1 - confidence_level) / 2, df)  # t-critical value
margin_of_error <- t_critical * standard_error

# Calculate the confidence interval
lower_bound <- sample_proportion - margin_of_error
upper_bound <- sample_proportion + margin_of_error

# Print the results
cat("Sample Proportion:", sample_proportion, "\n")
cat("95% Confidence Interval :", lower_bound, "to", upper_bound, "\n")

x <- seq(sample_proportion - 4 * standard_error, sample_proportion + 4 * standard_error, length = 1000)


y <- dt((x - sample_proportion) / standard_error, df) 

# Create a data frame for plotting
plot_data <- data.frame(x = x, y = y)



ggplot(plot_data, aes(x = x, y = y)) +
  geom_line() +
  geom_vline(xintercept = c(lower_bound, upper_bound), color = "blue", linetype = "dashed", size = 1) +
  geom_ribbon(data = plot_data[plot_data$x >= lower_bound & plot_data$x <= upper_bound, ], aes(x = x, ymin = 0, ymax = y), fill = "pink", alpha = 0.5) +
  labs(
    title = "Sample Distribution for the propotion of game in the Shooter Genre with a rating >=8.0",
    x = "Proportion of Shooter Games with User Score > 8.0",
    y = "Probability Density"
  )+
  theme_minimal()

```

As shown, the sample proportion is around ```r sample_proportion``` which implies that 31.6% of 'Shooter' games in our dataset had a user rating of >=8.0

The confidence The 95% confidence interval for this proportion falls between 0.283 to 0.350 . This indicates that the true proportion of highly-rated shooter games likely lies within this range

Interestingly, the proportion of of non-shooter games is around 0.337 which is higher than shooter games.

## Research Question 2 : 

**What are the average user scores for games developed by Nintendo, and how do they compare to other companies? Are these differences statistically significant?**

The objective of this research question was to look at games developed by the developer giant Nintendo, to determine is they did better on average compares to non-nintendo games, specifically in terms of their user scores and critic scores.

We define the Hypotheses as follows:

* Null Hypothesis : There is no difference between the mean scores of games developed by Nintendo and others
* Alternate Hypothesis : There is a non-zero difference between the mean scores of games developed by Nintendo and others

```{r}
#------------------------------------------------------------------------------------------------------------------
# SMART Q - Nintendo vs Non- Nintendo


dev_nintendo=subset(games_final,games_final$Developer=='Nintendo')

dev_others=subset(games_final,games_final$Developer!='Nintendo')

#USer scores

t_test_nintendo <- t.test(dev_nintendo$User_Score,dev_others$User_Score,var.equal = FALSE,conf.level=0.99) 

t_test_nintendo

#Critic scores

t_test_nintendo <- t.test(dev_nintendo$Critic_Score,dev_others$Critic_Score,var.equal = FALSE,conf.level=0.99) 

t_test_nintendo


```

The results of our 2 sample t-tests give us enough to reject the null hypothesis in both cases. Nintendo performs much better compared to other developers, having on average a user score of **7.94** compared to **7.13** and critic score of **77.16** compared to **68.95**

```{r}

#--- Visually showing results


ggplot(games_final, aes(x = ifelse(grepl("Nintendo", games_final$Developer), "Nintendo", "Other"), y = User_Score, color = ifelse(grepl("Nintendo", games_final$Developer), "Nintendo", "Other"))) +
  geom_point() +
  labs(
    title = "User Score for Nintendo Games vs. Others",
    x = "Developer",
    y = "User Score"
  ) +
  scale_color_manual(name = "Developer", values = c("Nintendo" = "skyblue", "Other" = "orange")) +
  scale_x_discrete(labels = c("Nintendo" = "Nintendo", "Other" = "Other")) +
  theme_minimal() +
  theme(legend.position = "top")
#----



ggplot(games_final, aes(x = ifelse(grepl("Nintendo", games_final$Developer), "Nintendo", "Other"), y = Critic_Score, color = ifelse(grepl("Nintendo", games_final$Developer), "Nintendo", "Other"))) +
  geom_point() +
  labs(
    title = "Critic Score for Nintendo Games vs. Others",
    x = "Developer",
    y = "Critic Score"
  ) +
  scale_color_manual(name = "Developer", values = c("Nintendo" = "skyblue", "Other" = "orange")) +
  scale_x_discrete(labels = c("Nintendo" = "Nintendo", "Other" = "Other")) +
  theme_minimal() +
  theme(legend.position = "top")
#----

```

The scatter plots also clearly illustrate this difference. Nintendo benifits from producing a much lower number of badly rated games, and they maintain consistency. Perhaps, this is the main reason they have maintained their brand image over the years as an elite game developer





